Underdispersion in the context of community athletics

### 2024-04-05 inst/parkrun_idea.Rmd
cite underdispersion
bugfix: there was something wrong with read.table()

@@

```{r calculateLikemat,cache=TRUE}

alpha_try <- seq(from=1,to=6,len=10)
beta_try <- 1:26

M <- as.matrix(expand.grid(alpha=alpha_try, beta=beta_try))
system.time(L <- t(apply(M,1,getlike)))
L_integer <- L  # counts
L[,1:26] <- L[,1:26]/mean(rowSums(L[,1:26]))
colnames(L) <- c(letters,'alpha','beta')
head(L)
```

@@

```{r}
plot(table(L_integer[,1:26]))
```

## Now 10 in-silico races:
```{r doinsilico,cache=FALSE}
alphatrue <- 1.3
betatrue <- 20   # 13th letter = "m"
set.seed(1)
D <- replicate(10,which(rrace(setNames(f(26,alphatrue),letters))==letters[betatrue]))
sort(D)
D <- 10:20
D
```

>>>>

@paul1995 state that "underdispersion will be seen very rarely in
real-life situations".

@@

```{r calculateLikemat,cache=FALSE}
alpha_try <- seq(from=1,to=6,len=20)
beta_try <- 1:26
calculate_from_scratch <- FALSE
if(calculate_from_scratch){
    M <- as.matrix(expand.grid(alpha=alpha_try, beta=beta_try))
    system.time(L <- t(apply(M,1,getlike)))
    colnames(L) <- c(letters,'alpha','beta')
    saveRDS(L,file="likelihoodmatrix.Rdata")
} else {
    L <- readRDS("likelihoodmatrix.Rdata")
}
L_integer <- L  # counts
L[,1:26] <- L[,1:26]/mean(rowSums(L[,1:26]))
head(L_integer)
head(L)
```

@@

```{r}
#plot(table(L_integer[,1:26]))
```

## Now 10 in-silico races:
```{r doinsilico,cache=FALSE}
alphatrue <- 2.3
betatrue <- 11 
set.seed(1)
D <- replicate(17,which(rrace(setNames(f(26,alphatrue),letters))==letters[betatrue]))
sort(D)
plot(table(D))
```
